Iterative Monomorphisation

### 2024-07-08 iterative_monomorphisation.tex
more
concl

\begin{abstract}
Monomorphisation can be used to extend monomorphic provers to support polymorphic logics. We propose an iterative approach, which is necessarily incomplete but which works well in practice. It is implemented in the Zipperposition prover, where it can be used to translate away polymorphism before invoking the monomorphic prover E as a backend. Our evaluation demonstrates that this approach increases Zipperposition's success rate. Moreover, we find iterative monomorphisation to be competitive with native implementations of polymorphism.
\end{abstract}

@@

In this paper, we present an algorithm based on our understanding of his description and implementation (Section~\ref{sec:high level-algorithm}). We also provide a more detailed description to help future implementers. In addition, this description shows some of the ways in which an implementation can be made more efficient and avoid explosions (Section~\ref{sec:low-level-algorithm}).

@@

% summary

% Future avenues

@@

* Sascha Böhme

>>>>

\begin{abstract}
Monomorphisation can be used to extend monomorphic provers to support polymorphic logics. We propose an iterative approach, which is necessarily incomplete but which works well in practice. It is implemented in the Zipperposition prover, where it can be used to translate away polymorphism before invoking the monomorphic prover E as a backend. Our evaluation demonstrates that this approach increases Zipperposition's success rate. Moreover, we find that iterative monomorphisation outperforms native implementations of polymorphism.
\end{abstract}

@@

In this paper, we present an algorithm based on our understanding of Böhme's description and implementation (Section~\ref{sec:high level-algorithm}). We also provide a more detailed description to help future implementers. In addition, this description shows some of the ways in which an implementation can be made more efficient and avoid explosions (Section~\ref{sec:low-level-algorithm}).

@@

We described a procedure for iteratively instantiating polymorphic types to
produce monomorphic problems as a preprocessor. Our primary motivation was to
improve the success rate of Zipperposition and its monomorphic E backend, and
indeed our evaluation shows a clear improvement. We also saw that even with
automatic provers that support polymorphism, iterative monomorphisation is a
better alternative in practice.

We see the following avenues for future work. First, iterative monomorphisation
blindly enumerates candidate instantiations, without exploiting any knowledge
about the logical structure of the formulas in which symbols occur. For
example, a lemma $\sym{p}\langle\alpha\rangle$ cannot be used to prove the
conjecture $\lnot \sym{p}\langle\ty{nat}\rangle$ because of the incompatible
polarities, but our procedure instantiates $\alpha$ with $\ty{nat}$ regardless.
Second, some automatic provers as well as tools like Sledgehammer include
a relevance filter that heuristically select a subset of the available axioms;
filters such as SInE \cite{xxx} and MePo \cite{xxx} are iterative and could be
interleaved with monomorphisation. Third, although one would expect native
implementations of polymorphism to outperform any preprocessor, our evaluation
clearly shows that this is not the case, therefore indicating that there is
considerable room for improvement on the native front.

@@

We thank Sascha Böhme for fruitful discussions.

This research is co-funded by the European Union (ERC, Nekoka, 101083038).
Views and opinions expressed are however those of the authors only and do not
necessarily reflect those of the European Union or the European Research
Council. Neither the European Union nor the granting authority can be held
responsible for them.
